---
title: "Duplex Seq Targeted Panel Design"
author: "Nate D. Olson"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

# Objective
Identify variants for targeted duplex sequencing panel for validating candidate mosaic variants.

# Background
Duplex sequencing uses dual molecular identifiers for strand specific sequencing of individual molecules, providing a theoretical error rate of $10^-9$ (Salk et al. 2018, https://doi.org/10.1038/nrg.2017.117).

This is arguably overkill for detecting our targeted 5% mutation threshold but will provide strong evidence for any validated variants.

Working with Twist to generate probes for the targeted panel (https://www.twistbioscience.com/) and Twinstrand duplex sequencing (https://twinstrandbio.com). Sequencing will be performed by CHOP sequencing center.

The targeted panel can include up to 5,000 probes, with 1 probe per variant.

# Panel Experimental Design 
3,000 probes

* 1,600 putative mosaic variants  
* 800 potential mosaic variants,
* 300 TP 
* 300 FP (homozygous reference HG002, heterozygous variant in parents).

__Mosaic Variants__
Variants identified with the Strelka2 somatic variant caller with HG002 seq data provided as tumor and the combined parents provided as normal. High coverage (~ 300X) Illumina data was used in the analysis.
The resulting variant call set was filtered to exclude variants outside the union of the AJtrio benchmark regions and complex variants. [TODO - Add definition for complex variants]

* Putative mosaic variants - passing variants
* Potential mosaic variants - non-passing variant, 

__Controls__  

* True positives (TP) - heterozygous variant in HG002 that are homozygous ref in HG003 – these would be used to validate assay in conjunction with DNA mixtures.  
* False positives (FP) - homozygous reference HG002 and heterozygous variant in parents  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

# Identification of Candidate Mosaic Variants
Using rtg-tools vcfeval

Downloading HG002 GIAB V4.1 benchmark variant and bed files
```
wget ftp://ftp-trace.ncbi.nlm.nih.gov/ReferenceSamples/giab/release/AshkenazimTrio/HG002_NA24385_son/NISTv4.1/GRCh37/HG002_GRCh37_1_22_v4.1_draft_benchmark.bed.gz
wget ftp://ftp-trace.ncbi.nlm.nih.gov/ReferenceSamples/giab/release/AshkenazimTrio/HG002_NA24385_son/NISTv4.1/GRCh37/HG002_GRCh37_1_22_v4.1_draft_benchmark.bed.gz.tbi 
wget ftp://ftp-trace.ncbi.nlm.nih.gov/ReferenceSamples/giab/release/AshkenazimTrio/HG002_NA24385_son/NISTv4.1/GRCh37/HG002_GRCh37_1_22_v4.1_draft_benchmark.vcf.gz
wget ftp://ftp-trace.ncbi.nlm.nih.gov/ReferenceSamples/giab/release/AshkenazimTrio/HG002_NA24385_son/NISTv4.1/GRCh37/HG002_GRCh37_1_22_v4.1_draft_benchmark.vcf.gz.tbi
```

Download hg37d5 sdf pre-formatted file
```
wget https://s3.amazonaws.com/rtg-datasets/references/1000g_v37_phase2.sdf.zip
```

vcfeval ran on Nate's 13 inch macbook pro laptop
vcfeval started at 9:45 AM 10/6
```
BENCHPATH=resources/hg002_benchmark
rtg vcfeval \
    -b ${BENCHPATH}/HG002_GRCh37_1_22_v4.1_draft_benchmark.vcf.gz \
    -c data/strelka2_wgs_parents-vs-child/HG002_hs37d5_300X_wg.vcf.gz \
    --bed-regions ${BENCHPATH}/HG002_GRCh37_1_22_v4.1_draft_benchmark.bed.gz \
    -o data/strelka2_wgs_parents-vs-child/vcfeval_comparison \
    -t resources/1000g_v37_phase2.sdf  \
    --all-records \
    --ref-overlap 
```

Next Step - intersect with AJ trio benchmark region and exclude SVs and complex variants

# Mosaic Variant List
```{r}
fp_vcf <- read_tsv(here("data","strelka2_wgs_parents-vs-child",
                        "vcfeval_comparison","fp.vcf.gz"),comment = "##") %>% 
    rename(CHROM = `#CHROM`)
fp_vcf %>% 
    count(FILTER)
```
```{r}

```


# Control Variant List

# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```


## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```