---
title: "Seq Experimental Design"
author: "Nate Olson"
date: "1/6/2021"
output: html_document
---

```{r}
library(tidyverse)
library(here)
```

# Objective
Work out experimental design for duplex seq mosaic variant validation

# Approach

1. Use binomial probability to get consensus duplex seq coverage requirements. 
2. Get flow cell/ sample coverage using Illumina coverage calculator
3. Work out experimental design based on flow cell/ sample info

## Assumptions

### Coverage calculator
- duplication rate: 10% (based on email from Renata)
- on-target: 65% (based on email from Renata)
- coverage: 100X sample size value (based on conversation with TwinStrand)
- region_size: 5Mb [100bp * 5000 probes (2 per variant) ]

# Analysis

Number of consensus duplex seq coverage
```{r}
sample_size <- seq(from = 10, to = 5000, by = 1)
probs <- seq(from = 0.00, to = 0.1, by = 0.001)

detect_prob_tbl <- tibble(dup_cov = rep(sample_size, length(probs)),
                          af = rep(probs, each = length(sample_size))) %>% 
    mutate(prob_detect = 1 - pbinom(q = 0, dup_cov, af))
```


```{r}
thresh_vals <- detect_prob_tbl %>% 
    filter(af %in% c(0.001, 0.01, 0.05, 0.1)) %>% 
    filter(prob_detect > 0.95) %>% 
    group_by(af) %>% 
    top_n(n = 1, wt = -dup_cov)

detect_prob_tbl %>% 
     filter(af %in% c(0.001, 0.01, 0.05, 0.1)) %>% 
    ggplot() + 
    geom_line(aes(x = dup_cov, y = prob_detect, group = factor(af)), size = 1, color = "grey40") + 
    geom_line(aes(x = dup_cov, y = prob_detect, color = factor(af))) + 
    geom_point(data = thresh_vals, 
               aes(x = dup_cov, y = prob_detect, fill = factor(af)), 
               shape = 21) + 
    geom_linerange(data = thresh_vals, 
                   aes(x = dup_cov, ymin = 0, ymax = prob_detect), 
                   color = "grey60", linetype = 2) +
    scale_x_log10() + 
    theme_bw() + 
    labs(x = "Duplex Consensus Seq Coverage", y = "Probability of Detecting Variant", fill = "AF")
```
```{r}
detect_prob_tbl %>% 
    filter(prob_detect > 0.95) %>% 
    group_by(af) %>% 
    top_n(n = 1, wt = -dup_cov)
```
For AF of 0.01 looking at 300X duplex seq cov
For AF of 0.05 60X cov

For raw seq coverage multiply by 100 (I think, need to verify assumption), so between 6 and 30k X coverage.

## Illumina Coverage calculator
From Illumina coverage calculator webapp


    Calculation of Results Using Coverage Needed
    Total output required = region size * coverage / ((1-duplicates/100) * on target/100)
    
    Output per unit = clusters per unit * read length
    
    Number of units (flow cells or lanes) = total output required / output per unit
    
    Number of samples = output per unit / total output required


Flowcell sizes from illumina
```{r}
read_csv(here("data", "panel_design", "CoverageNeeded_20210106082807_02699900.csv"), 
              skip = 6, n_max = 6) 
```

## Read length based coverage calculation

```{r}
region_size <- 5000000 #5MB
duplicates <- 0.1
on_target <- 0.65

## Illumina NovaSeq run sizes
flowcell_output <- tibble(flowcell = c("SP", "S1", "S2", "S4"),
                              output_per_unit = c(160000000000, 
                                                  320000000000,
                                                  820000000000,
                                                  2000000000000))

## Cov Calc
tibble(dup_cov = c(50, 100, 150, 200, 250, 300, 400, 500),
       ill_flowcell = list(flowcell_output)) %>% 
    mutate(required_output = region_size * dup_cov * 100 / ((1-duplicates) * on_target)) %>% 
    unnest() %>% 
    mutate(num_samples = floor(output_per_unit/ required_output)) %>% 
    arrange(-num_samples) %>% 
    select(dup_cov, flowcell, num_samples) %>% 
    spread(flowcell, num_samples)
```

## Read count based estimate

```{r}
## Accounting for duplication and on target rates
panel_size <- 5000

tibble(
    flowcell = c("SP", "S1", "S2", "S4"),
    cluster_per_unit = c(800000000,
                        1600000000,
                         4100000000,
                        10000000000)) %>% 
    mutate(clusters_per_sample = cluster_per_unit/25,
           probe_cov = (clusters_per_sample * (1-duplicates) * on_target)/panel_size,
           dup_cov = floor(probe_cov/100)) 
```

AF detection thresholds for full S2 and S4 runs based on cluster numbers. 

```{r}
detect_prob_tbl %>% 
    filter(prob_detect > 0.95) %>% 
    group_by(af) %>% 
    top_n(n = 1, wt = -dup_cov) %>% 
    filter(dup_cov %in% c(199, 427))
```

# Experimental Design
Twin - reagents for 24 samples

Samples:
- Unmixed: HG002, HG003, HG004
- Mixed HG002/HG003: AF 0.25, 0.1, 0.05, 0.01

Triplicates for each

Estimated Dup Coverage:
S2 - 199 per replicate, ~600X total
S4 - 427 per replicate, ~1275X total

# Summary
AF detection limits based on the above estimates indicate that we can detect mosaic variants with AF ~0.015 with S2 flowcell and 0.007 using a S4 flowcell. 
The coverage calculations are pretty far off between the cluster based and base coveraer estimates.

__Caveats__
+ Sample size calculation does not include a power analysis  
+ Coverage estimate assumptions need verification, particularly the number of raw reads required per duplex consensus seq. 